The Controlled Choice Design and Paternalism in Pawnshop Borrowing

### 2023-10-15 donde.tex
revision after jpe submission

We define a borrower's financial cost as the total monetary outflow --in cash or pawn value-- from the borrower to the lender. This includes all payments the borrower made toward interest and fees, but also the value of the pawn when there is default. When there is no default the borrower gets her pawn back and there is no loss of value for the borrower. Payments towards capital are considered a cost only when the borrower defaults, as she does not get reimbursed for these. Note however that when she does not default payments to capital are not an actual outflow, as they sum up to the value of the loan the lender disbursed in the first place. The formula for financial cost for person $i$ is thus as follows:

\begin{align*}
    \text{Financial Cost}_i =&  \underbrace{\sum_t P^i_{it}}_{\text{Payments to interest}} + \underbrace{\sum_t P^f_{it}}_{\text{Payments of fees}}   \\
    &\quad\qquad + \underbrace{\mathds{1}(\text{Default}_i) \times (\text{Appraised Value}_i}_{\text{Cost of losing pawn}} + \underbrace{\sum_t P^c_{it}}_{\text{Payments to capital}})
\end{align*}

@@

Figure \ref{fan_park_bounds} plots the sharp bounds for $F_{\delta}$ based on the APR outcome defined above in Section \ref{costs}. Our point estimates of $\underline{F}(0)$ and $\overline{F}_0$ are .03 and .76 respectively, with associated 95\% confidence intervals of [0.027, 0.040] and [0.73, 0.78], which says that at least 24\% of individual borrowers benefit from commitment (and at most 97\%).\footnote{Confidence intervals are constructed using the asymptotic distribution for the bounds. See \cite{fan2010sharp} for details.}

@@

\begin{align*}
    \underbrace{\text{Financial Cost}_i}_{(1)} =&  \underbrace{\sum_t P^i_{it}}_{\underset{(2)}{\text{Payments to interests}}}  +\underbrace{\sum_t P^f_{it}}_{\underset{(3)}{\text{Payments of fees}}} + \underbrace{\overbrace{\mathds{1}(\text{Default}_i)}^{(6)}\times (\text{Appraised Value}_i}_{\underset{(5)}{\text{Cost of losing pawn}}} + \underbrace{\sum_t P^c_{it}}_{\underset{(4)}{\text{Payments to capital}}})
\end{align*}

@@

\begin{table}[H]
\caption{Type I \& II errors using targeting narrow rules}
\label{hit_miss_rule}
\begin{center}
\resizebox{0.95\textwidth}{!}{
\small{\input{./Tables/hit_miss_rule.tex}}
}
\end{center}
\scriptsize This table reports error rates for six different rules for allocating individuals to commitment. The first row allocates all borrowers to control. Taking the conditional ATE estimates from the ``wide'' causal forest, described in Section \ref{sec:RF} as ground truth, this results in 87\% of individuals losing by not receiving commitment. The second row assigns all borrowers to forced commitment, yielding error rates that are the exact mirror image of those from the first row.  The third row considers the infeasible optimum in which each borrower is allocated the ``correct'' treatment, treating the estimate from the ``wide'' causal forest as ground truth. The fourth and fifth rows assign borrowers to commitment based on a classification model that uses the narrow covariate set to predict an indicator for whether the ``wide'' random forest CATE estimate is positive. Row four uses a random forest classification model while row five uses a simple logistic regression model. In each of these rows, the assignment rule ranks borrowers from highest to lowest by their estimated probability of having a positive CATE. The 91\% of borrowers with the highest estimated probabilities are assigned to treatment, matching the overall rate of positive CATE estimates from the ``wide'' RF model. The sixth row gives the percentage of miss-classification that would follow from the observed choice. 
%assigns the commitment contract to all borrowers with a positive conditional ATE estimate from the ``narrow'' RF. The fifth row uses a logistic regression to estimate the probability that a given borrower has a positive conditional ATE, assigning commitment to those with an estimated \hl{probability greater than 0.5}.

@@

\begin{figure}[H]
    \caption{Partition of TUT by behavioral variables.}
    \label{tut_beh_partition}
    \begin{center}
        \centering
        \includegraphics[width=0.75\textwidth]{Figuras/tut_beh_partition.pdf} 
    \end{center}
     \scriptsize  Each panel in this figure shows how the estimated treatment on the untreated (TUT) effect varies with a binary survey variable $X_i$. In the left panel (P.B.), $X_i = 1$ if borrower $i$ is ``present-biased'' based on her responses to the time preference questions from our survey. In the right panel (Sure-confidence) $X_i = 1$ if  borrower $i$ reported that she was certain to recover her pawn, zero otherwise. Each panel depicts three estimates and associated 95\% confidence bands: the overall TUT (blue), the TUT for borrowers with $X = 1$ (green), and the TUT for borrowers with $X = 0$ (red). By the law of iterated expectations, the overall TUT in each panel must lie between the group-specific TUTs. The overall TUT is slightly different in each panel, because some borrowers only completed part of the survey. In each panel, we present results for the full set of borrowers who answered the survey question or questions needed to determine $X_i$. From the left panel, we see that TUT does not appear to vary with present bias. In contrast, the right panel suggests that the positive and significant overall TUT is driven by individuals who are ``sure-confident,'' i.e.\ borrowers who report that they are certain to recover their pawn.
     %to be largely concentrated relevant survey question  according to her survey responses two binary performs a partition of the Treatment on the Untreated (FC), estimating the share of the effect that arises within each of two binary categories.  The top bars examine present bias (individuals are more impatient over one-month time horizon than they are between three and four months).  The bottom bars examine sure confidence (individual reports 100\% probability of repaying at time of taking loan).  Blue bar is the component arising from those without the trait, orange bar from those with the trait. 
      %\footnotesize{ \textit{Do file: }  \texttt{partition_tut.do}

@@

  %  \end{center}
        %\scriptsize
        %In Panel (a) we estimate what the treatment effect of the commitment contract \textit{would have been} for all subjects in the choice arm if they had been forced into the fee-commitment contract. To do this we proceed in two steps. First, we estimate treatment effects in the forcing arm by comparing the Forced Commitment arm against the status quo arm. We let these effects be heterogeneous as a function of our $x$'s using \cite{atheygrf}'s methodology of causal forests. Second, we extrapolate these treatment effects based on the choice arm using the same $x$'s. Once we have personalized counterfactual treatment effects in the choice arm we calculate what fraction of subjects in the choice arm incurred in financial costs that are  $> z$\% than if they had chosen the opposite contract of what they actually chose, were $z$\% is defined as a fraction of the loan. $z$\% is a level of tolerance we can vary and we plot it in the X-axis. The left Y-axis measures the fraction of subjects that would have been better by a margin of $z$\% if they changed their choice, and the right Y-axis measures the amount of money ``left on the table''. We use bootstrap to tighten the confidence intervals (CIs). \cite{atheygrf}'s heterogeneous treatment effect using GRF is asymptotically normal. We compute the HTE ($\mu$) together with standard errors ($\sigma$) . For every pledge, we draw a random effect from a normal distribution with parameters ($\mu$,$\sigma^2$), and compute via bootstrap the percentage that choose wrong, along normal-approximation CIs (results are robust if we use instead percentile CIs or bias-corrected CIs) . This allows us to estimate a distribution for the upper and lower bound CIs, which we then use to obtain a 95\% CI. Panel (b) is analogous to Panel (a) except that it does the exercise separately for clients we classified as overconfident ($OC_i:=\mathbbm{1}(P^s_i-\widehat{P_i(X_i)}>0))$ and those we classified as not overconfident. Panel (c) simulates what percentage would be better by at least $z$ if we forced everybody of those in the Commitment Choice arm into the fee-commitment contract.
        % Panel (d) asks whether it is the case that people with larger causal benefits from the Forced Commitment contract would be more likely to select that contract. To do this we proceed in two steps as well. First we estimate a flexible model of take-up of the fee commitment contract \textit{in the choice arm} using random forests, and from this model obtain a probability $P(x_i)$ of choosing the Forced Commitment contract for a client with characteristics $x_i$. We extrapolate this model to clients in the Forced Commitment arm to estimate what would they have chosen if we had given them choice. 
        % Figure (d) plots $\widehat{P(x_i)}$ in the X-axis using a binscatter that splits the X-axis in 100 percentile bins. The Y-axis plots the heterogeneous treatment effects of the fee forcing contract on the probability of losing the pawns (more negative means then more likely to recover it), averaged for the respective x-axis bin. Positive assortative selection would mean a negative relationship: those who benefit more by treatment and lose the pawn less are \emph{less} likely to choose it.  Using all bins we estimate a slope of zero in the relationship. Using the 80 right most points, we estimate a \textit{positive} relationship.
        %\textit{Do file: }  \texttt{choose\_wrong\_quant\_wrong.do, choose\_wrong\_quant\_wrong\_decomposition.do}
\end{figure}

@@

\begin{figure}[H]
    \centering
  \caption{Cumulative Distribution Function of Conditional ATE Estimates}
  \includegraphics[width=0.75\textwidth]{Figuras/line_better_forceall_apr_te_cf.pdf} 
    \label{fig:CATEsurvival}
\end{figure}

This figure shows the share of borrowers who benefit from forced commitment, where ``benefit'' is defined as having an estimated conditional average treatment effect (ATE) above a specified threshold value. Results are based on the causal forest model from Section \ref{sec:RF}. The solid line equals $1 - F_\text{ATE}(\delta)$ where $F_\text{ATE}(\delta)$ denotes the CDF corresponding to the density of conditional ATE estimates from Figure \ref{heterogeneous_effects}. The blue shaded regions are associated 95\% confidence intervals. We estimate that at least 91\% of borrowers have a positive conditional ATE.

>>>>

We define a borrower's financial cost as the total monetary outflow --in cash or pawn value-- from the borrower to the lender. This includes all payments the borrower made toward interest and fees, but also the net difference of the appraised value of the loan and the loan amount (VLD) when there is default. When there is no default the borrower gets her pawn back and there is no loss of value for the borrower. Payments towards capital are considered a cost only when the borrower defaults, as she does not get reimbursed for these. Note however that when she does not default payments to capital are not an actual outflow, as they sum up to the value of the loan the lender disbursed in the first place. The formula for financial cost for person $i$ is thus as follows\footnote{We can also define the financial cost formula as:
\begin{align*}
    \text{Financial Cost}_i =&  \underbrace{\sum_t P^i_{it}}_{\text{Payments to interest}} + \underbrace{\sum_t P^f_{it}}_{\text{Payments of fees}}    + \underbrace{\sum_t P^c_{it}}_{\text{Payments to capital}} + \underbrace{\mathds{1}(\text{Default}_i) \times (\text{Appraised value}_i}_{\text{Cost of losing pawn}})-\text{Loan amount}_i
\end{align*}
and our results will be equivalent if using this definition instead.
}:

\begin{align*}
    \text{Financial Cost}_i =&  \underbrace{\sum_t P^i_{it}}_{\text{Payments to interest}} + \underbrace{\sum_t P^f_{it}}_{\text{Payments of fees}}   \\
    &\quad\qquad + \underbrace{\mathds{1}(\text{Default}_i) \times (\text{Val-Loan Difference}_i}_{\text{Cost of losing pawn}} + \underbrace{\sum_t P^c_{it}}_{\text{Payments to capital}})
\end{align*}

@@

Figure \ref{fan_park_bounds} plots the sharp bounds for $F_{\delta}$ based on the APR outcome defined above in Section \ref{costs}. Our point estimates of $\underline{F}(0)$ and $\overline{F}_0$ are .03 and .77 respectively, with associated 95\% confidence intervals of [0.024, 0.050] and [0.75, 0.80], which says that at least 23\% of individual borrowers benefit from commitment (and at most 97\%).\footnote{Confidence intervals are constructed using the asymptotic distribution for the bounds. See \cite{fan2010sharp} for details.}

@@

\begin{align*}
    \underbrace{\text{Financial Cost}_i}_{(1)} =&  \underbrace{\sum_t P^i_{it}}_{\underset{(2)}{\text{Payments to interests}}}  +\underbrace{\sum_t P^f_{it}}_{\underset{(3)}{\text{Payments of fees}}} + \underbrace{\overbrace{\mathds{1}(\text{Default}_i)}^{(6)}\times (\text{Val-Loan Diff}_i}_{\underset{(5)}{\text{Cost of losing pawn}}} + \underbrace{\sum_t P^c_{it}}_{\underset{(4)}{\text{Payments to capital}}})
\end{align*}

@@

\begin{table}[H]
\caption{Type I \& II errors using targeting narrow rules}
\label{hit_miss_rule}
\begin{center}
\resizebox{0.95\textwidth}{!}{
\small{\input{./Tables/hit_miss_rule.tex}}
}
\end{center}
\scriptsize This table reports error rates for six different rules for allocating individuals to commitment. The first row allocates all borrowers to control. Taking the conditional ATE estimates from the ``wide'' causal forest, described in Section \ref{sec:RF} as ground truth, this results in 93\% of individuals losing by not receiving commitment. The second row assigns all borrowers to forced commitment, yielding error rates that are the exact mirror image of those from the first row.  The third row considers the infeasible optimum in which each borrower is allocated the ``correct'' treatment, treating the estimate from the ``wide'' causal forest as ground truth. The fourth and fifth rows assign borrowers to commitment based on a classification model that uses the narrow covariate set to predict an indicator for whether the ``wide'' random forest CATE estimate is positive. Row four uses a random forest classification model while row five uses a simple logistic regression model. In each of these rows, the assignment rule ranks borrowers from highest to lowest by their estimated probability of having a positive CATE. The 89\% of borrowers with the highest estimated probabilities are assigned to treatment, matching the overall rate of positive CATE estimates from the ``wide'' RF model. The sixth row gives the percentage of miss-classification that would follow from the observed choice. 
%assigns the commitment contract to all borrowers with a positive conditional ATE estimate from the ``narrow'' RF. The fifth row uses a logistic regression to estimate the probability that a given borrower has a positive conditional ATE, assigning commitment to those with an estimated \hl{probability greater than 0.5}.

@@

\begin{figure}[H]
    \caption{Partition of TUT by behavioral variables}
    \label{tut_beh_partition}
    \begin{center}
        \centering
        \includegraphics[width=0.75\textwidth]{Figuras/tut_beh_partition.pdf} 
    \end{center}
     \scriptsize  Each panel in this figure shows how the estimated treatment on the untreated (TUT) effect varies with a binary survey variable $X_i$. In the left panel (P.B.), $X_i = 1$ if borrower $i$ is ``present-biased'' based on her responses to the time preference questions from our survey. In the right panel (Sure-confidence) $X_i = 1$ if  borrower $i$ reported that she was certain to recover her pawn, zero otherwise. Each panel depicts three estimates and associated 95\% confidence bands: the overall TUT (blue), the TUT for borrowers with $X = 1$ (green), and the TUT for borrowers with $X = 0$ (red). By the law of iterated expectations, the overall TUT in each panel must lie between the group-specific TUTs. The overall TUT is slightly different in each panel, because some borrowers only completed part of the survey. In each panel, we present results for the full set of borrowers who answered the survey question or questions needed to determine $X_i$. From the left panel, we see that TUT does not appear to vary with present bias. In contrast, the right panel suggests that the positive and significant overall TUT is driven by individuals who are ``sure-confident,'' i.e.\ borrowers who report that they are certain to recover their pawn.
     %to be largely concentrated relevant survey question  according to her survey responses two binary performs a partition of the Treatment on the Untreated (FC), estimating the share of the effect that arises within each of two binary categories.  The top bars examine present bias (individuals are more impatient over one-month time horizon than they are between three and four months).  The bottom bars examine sure confidence (individual reports 100\% probability of repaying at time of taking loan).  Blue bar is the component arising from those without the trait, orange bar from those with the trait. 
      %\footnotesize{ \textit{Do file: }  \texttt{partition_tut.do}

@@

  %  \end{center}
        %\scriptsize
        %In Panel (a) we estimate what the treatment effect of the commitment contract \textit{would have been} for all subjects in the choice arm if they had been forced into the fee-commitment contract. To do this we proceed in two steps. First, we estimate treatment effects in the forcing arm by comparing the Forced Commitment arm against the status quo arm. We let these effects be heterogeneous as a function of our $x$'s using \cite{atheygrf}'s methodology of causal forests. Second, we extrapolate these treatment effects based on the choice arm using the same $x$'s. Once we have personalized counterfactual treatment effects in the choice arm we calculate what fraction of subjects in the choice arm incurred in financial costs that are  $> z$\% than if they had chosen the opposite contract of what they actually chose, were $z$\% is defined as a fraction of the loan. $z$\% is a level of tolerance we can vary and we plot it in the X-axis. The left Y-axis measures the fraction of subjects that would have been better by a margin of $z$\% if they changed their choice, and the right Y-axis measures the amount of money ``left on the table''. We use bootstrap to tighten the confidence intervals (CIs). \cite{atheygrf}'s heterogeneous treatment effect using GRF is asymptotically normal. We compute the HTE ($\mu$) together with standard errors ($\sigma$) . For every pledge, we draw a random effect from a normal distribution with parameters ($\mu$,$\sigma^2$), and compute via bootstrap the percentage that choose wrong, along normal-approximation CIs (results are robust if we use instead percentile CIs or bias-corrected CIs) . This allows us to estimate a distribution for the upper and lower bound CIs, which we then use to obtain a 95\% CI. Panel (b) is analogous to Panel (a) except that it does the exercise separately for clients we classified as overconfident ($OC_i:=\mathbbm{1}(P^s_i-\widehat{P_i(X_i)}>0))$ and those we classified as not overconfident. Panel (c) simulates what percentage would be better by at least $z$ if we forced everybody of those in the Commitment Choice arm into the fee-commitment contract.
        % Panel (d) asks whether it is the case that people with larger causal benefits from the Forced Commitment contract would be more likely to select that contract. To do this we proceed in two steps as well. First we estimate a flexible model of take-up of the fee commitment contract \textit{in the choice arm} using random forests, and from this model obtain a probability $P(x_i)$ of choosing the Forced Commitment contract for a client with characteristics $x_i$. We extrapolate this model to clients in the Forced Commitment arm to estimate what would they have chosen if we had given them choice. 
        % Figure (d) plots $\widehat{P(x_i)}$ in the X-axis using a binscatter that splits the X-axis in 100 percentile bins. The Y-axis plots the heterogeneous treatment effects of the fee forcing contract on the probability of losing the pawns (more negative means then more likely to recover it), averaged for the respective x-axis bin. Positive assortative selection would mean a negative relationship: those who benefit more by treatment and lose the pawn less are \emph{less} likely to choose it.  Using all bins we estimate a slope of zero in the relationship. Using the 80 right most points, we estimate a \textit{positive} relationship.
        %\textit{Do file: }  \texttt{choose\_wrong\_quant\_wrong\_tot\_tut.do, choose\_wrong\_quant\_wrong\_decomposition.do}
\end{figure}

@@

\begin{figure}[H]
    \centering
  \caption{Cumulative Distribution Function of Conditional ATE Estimates}
  \includegraphics[width=0.75\textwidth]{Figuras/cdf_CATE.pdf} 
    \label{fig:CATEsurvival}
\end{figure}

This figure shows the share of borrowers who benefit from forced commitment, where ``benefit'' is defined as having an estimated conditional average treatment effect (ATE) above a specified threshold value. Results are based on the causal forest model from Section \ref{sec:RF}. The solid line equals $1 - F_\text{ATE}(\delta)$ where $F_\text{ATE}(\delta)$ denotes the CDF corresponding to the density of conditional ATE estimates from Figure \ref{heterogeneous_effects}. The blue shaded regions are associated 95\% confidence intervals. We estimate that at least 91\% of borrowers have a positive conditional ATE. 
%\textit{Do file: }  \texttt{cdf_CATE.do}
